ScipionBox: automatic processing for

### 2017-07-24 1.introduction.tex

3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes. The field is under a major transformation due to the arrival of better microscopes [**REF], new large area detectors [**REFS ddd] and automatization [**REFs to data colection software as EPU, leginon, serialEm,etc...]  which makes possible that a single microscope generates high quality data sets on the order of terabytes per day [**REF] while working for several days without interruption. Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of experienced microscopist and; (3) the need for an effective transference of data and associated metadata in an increasingly distributed and collaborative environment. This situation places techniques that are automated and allow higher throughput in high demand.

@@

Accounting for major government investment in increasingly costly electron microscopes requires a large user  base. Clearly, the future of projects that require this high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff. \scipionbox has been designed with this scenario in mind in which the early detection of issues may save the data acquisition shift which otherwise need to be repeated after a new application that may be allowed many months later.

>>>>

3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes. The field is under a major transformation due to the arrival of better microscopes [**REF], new large area detectors [**REFS ddd] and automatization [**REFs to data colection software as EPU, leginon, serialEm,etc...]  which makes possible that a single microscope generates high quality data sets on the order of terabytes per day [**REF] while working for several days without interruption. Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective transference of data and associated metadata in an increasingly distributed and collaborative environment. This situation places techniques that are automated and allow higher throughput in high demand.

@@

Accounting for major government investment in increasingly costly electron microscopes requires a large user  base. Clearly, the future of projects that require this high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff. \scipionbox has been designed with this scenario in mind in which the early detection of issues may save the data acquisition shift which otherwise need to be repeated after a new application that may be granted many months later.

### 2017-07-24 2.overall.tex

\section{Overall System Design: Worflows, Protocols, Monitors, Consensus and Beyond}
\label{overall}
Just like snowflakes, no two \scipionbox installation are alike. So far, the system is running in 
three centralized facilities, and each installation has been tailored to the particular needs of each facility. The original design satisfied the requirements of our home facility (National Center for Biotechnology, Madrid, Spain), in which \scipionbox automatically fetches newly recorded movies from a network mounted disk. After this first installation, \scipionbox was deployed on XXXXX (a centralized facility located at Stockholm [**provide full identification @delarosa]) where a rigorous booking system is followed. There, \scipionbox was adapted to interface with this booking platform in order to produce better reports. Lately,  \scipionbox has been installed in a large Synchrotron  such as Diamond (Oxford, UK) where projects are handled by ISPyB (a customized lims -laboratory management information system- [**REF]) and data are saved on a distributed file system that heavily penalizes access to lists of files. In this environment, \scipionbox needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created. All these particularities, that will be described in detail later, could only be handled thanks to a careful design that prioritizes flexibility at many levels.

@@

In addition to offering algorithms (in Scipion lingo ``protocols'') we provide 
a set of ``monitors'', which are protocols that  check on the progress of other protocols. The monitors are designed to produce analysis plots, generate reports or raise alerts when some problems are detected. A monitor example is the CTF monitor, that checks the computed defocus values for
each micrograph and can raise an alert if the values are above or below certain thresholds.
Monitors were also designed in a modular way to be modified or extended. 
Monitors may send alert emails, continuously generate html files or interact with lims so that users and staff may easily follow the data acquisition and processing in-house or remotely. At the project end, a review of the processing history is produced by default.
%This feature was very important since we knew that difference facilities will have different needs and criteria to be fulfilled.

@@

\scipionbox is a streaming system that runs on threads. This is a significant difference from a script, since it means that each module is run independently. For example, Do you want to add a new protocol or monitor in your workflow while it is running? No problem, do it. \scipionbox allows users to modify  workflows on the fly without modifiying a single code line. The modification will not interfere with the already executed steps and may take advantage from all already computed data. Did something go wrong and the microscope stopped? By default \scipionbox will stop after 2 hours with no microscope activity, but if you wish to continue your acquisition after a stop just press the continue button [*+ scipion 1.1.2], there is no need to restart from the beginning.

>>>>

\section{Overall System Design: Worflows, Protocols, Monitors, Consensus and Beyond}
\label{overall}
Just like snowflakes, no two \scipionbox installation are alike. So far, the system is running in 
three centralized facilities, and each installation has been tailored to the particular needs of each facility. The original design satisfied the requirements of our home facility (National Center for Biotechnology, Madrid, Spain), in which \scipionbox automatically fetches newly recorded movies from a network mounted disk. After this first installation, \scipionbox was deployed on XXXXX (a centralized facility located at Stockholm [**provide full identification @delarosa]) where a rigorous booking system is followed. There, \scipionbox was adapted to interface with this booking platform in order to produce better reports. Lately,  \scipionbox has been installed in a large Synchrotron  such as Diamond (Oxford, UK) where projects are handled by ISPyB (a customized laboratory management information system -LIMBS- [**REF]) and data are saved on a distributed file system that heavily penalizes access to lists of files. In this environment, \scipionbox needs to constantly interchange information with ISPyB, and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created. All these particularities, that will be described in detail later, could only be handled thanks to a careful design that prioritizes flexibility at many levels.

@@

In addition to offering algorithms (in Scipion lingo ``protocols'') we provide 
a set of ``monitors'', which are protocols that  check on the progress of other protocols. The monitors are designed to produce analysis plots, generate reports or raise alerts when some problems are detected. A monitor example is the CTF monitor, that checks the computed defocus values for
each micrograph and can raise an alert if the values are above or below certain thresholds.
Monitors were also designed in a modular way to be modified or extended. 
Monitors may send alert emails, continuously generate HTML files or interact with LIMS so that users and staff may easily follow the data acquisition and processing in-house or remotely. At the project end, a review of the processing history is produced by default.
%This feature was very important since we knew that difference facilities will have different needs and criteria to be fulfilled.

@@

\scipionbox is a streaming system that runs on threads. This is a significant difference from a script, since it means that each module is run independently. For example, Do you want to add a new protocol or monitor in your workflow while it is running? No problem, do it. \scipionbox allows users to modify  workflows on the fly without modifying a single code line. The modification will not interfere with the already executed steps and may take advantage from all already computed data. Did something go wrong and the microscope stopped? By default \scipionbox will stop after 2 hours with no microscope activity, but if you wish to continue your acquisition after a stop just press the continue button [*+ scipion 1.1.2], there is no need to restart from the beginning.

### 2017-07-24 3.workflow.tex

@@

\subsection{CNB Facility}

@@

\subsection{XX Facility}

\subsection{YY Facility}

%cat /proc/cpuinfo
%\begin{figure}[h]
%\centering
%     \includegraphics[width=50mm]{Images/to-be-done}
%     \caption{TODO: DROP? I do not know if an image showing CNB setup is needed this image should show CNB setup, microscope connected to control-PC which saves data in \scipionbox using samba. Show Ethernet speed. Soy connection to external disk/host.
%     Describe number GPUs, CPUs, SSD disks.
%     }
%     \label{cnbsetup}
%\end{figure}

%One of the most frequent requests by both beginners, and even experienced digital darkroom workers, is for detailed step-by-step instructions on how to perform various basic tasks in Photoshop.

%One of the key goals of ScipionBox is to guarantee interoperability between algorithms implemented in different software packages.  The possibility of choosing among several algorithms is provided at all levels of the workflow, in many cases adding the possibility of comparing the results of these methods to detect inconsistencies.  
%and do not need to be concerned with how the rest of the workflow is allocated.

%ScipionBox design is agnostic to a particular workflow or software packet.

%%Export options?

>>>>

%Comment change in scale in the three examples: CNB, SNC and DIAMOND
%more local, SNC national, Diamond international. Portal infrastructure has
%different levels.

@@

\subsection{CNB Facility}
%Here there are Scipion general features and CNB specific. Move the general to the Processing with Scipion section.

@@

\subsection{Swedish National CryoEM Facility}

\begin{verbatim}
We need to cover here differences with CNB setup
 CNB is a stand alone center. It does not work as a real National Facility
 SNC already had an application portal and Scipion needs to interact
 with it. 
 - initial script
 - same as CNB until session end
 - when acquisistion done contaCT PORTAL AND UPLOAD REPORT
\end{verbatim}

\subsection{Diamond Light Source Facility??? or eBIC???}

\begin{verbatim}

We need to cover here differences with CNB setup

 DEVELOPER RELATED
 * Describe infrastructure in diamond: file system (distributed, user permissions),
                                       no graphical output, queues, 
 * Describe Scipion modification:      import (detect new files by socket message)
                                       monitor, report to ISPyB 
                                       scipion headless
                                       execution as script

USER RELATED
 * how users interact with scipion?
 * how data is visualized?
\end{verbatim}

%%%Diamond Light Source Facility  ends here%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%cat /proc/cpuinfo
%\begin{figure}[h]
%\centering
%     \includegraphics[width=50mm]{Images/to-be-done}
%     \caption{TODO: DROP? I do not know if an image showing CNB setup is needed this image should show CNB setup, microscope connected to control-PC which saves data in \scipionbox using samba. Show Ethernet speed. Soy connection to external disk/host.
%     Describe number GPUs, CPUs, SSD disks.
%     }
%     \label{cnbsetup}
%\end{figure}
